---
# IMPORTANT: Change settings here, but DO NOT change the spacing.
# Remove comments and add values where applicable.
# The descriptions below should be self-explanatory

title: "Question 3: Portfolio Construction"
#subtitle: "Hyperbolic discounting in new car purchases"

documentclass: "elsarticle"

# --------- Thesis title (Optional - set to FALSE by default).
# You can move the details below around as you please.
Thesis_FP: FALSE
# Entry1: "An unbelievable study with a title spanning multiple lines."
# Entry2: "\\textbf{Some Guy}" # textbf for bold
# Entry3: "A thesis submitted toward the degree of Doctor of Philosophy"
# Uni_Logo: Tex/Logo.png # Place a logo in the indicated location (from your root, e.g. defaults to ~/Tex/Logo.png) and uncomment this line. Leave uncommented for no image
# Logo_width: 0.3 # If using a logo - use this to set width (size) of image
# Entry4: "Under the supervision of: \\vfill Prof. Joe Smith and Dr. Frank Smith"
# Entry5: "Stellenbosch University"
# Entry6: April 2020
# Entry7:
# Entry8:

# --------- Front Page
# Comment: ----- Follow this pattern for up to 5 authors
AddTitle: TRUE # Use FALSE when submitting to peer reviewed platform. This will remove author names.
Author1: "Liam Andrew Beattie"  # First Author - note the thanks message displayed as an italic footnote of first page.
Ref1: "Financial Econometrics 871, Stellenbosch University, South Africa" # First Author's Affiliation
Email1: "22562435\\@sun.ac.za" # First Author's Email address

# Author2: "Abdul Qaadir Cassiem"
# #Ref2: "Some other Institution, Cape Town, South Africa"
# Email2: "20863667\\@sun.ac.za"
# CommonAffiliation_12: TRUE # If Author 1 and 2 have a common affiliation. Works with _13, _23, etc.

# Author3: "John Doe"
# Email3: "Joe\\@gmail.com"

#CorrespAuthor_1: TRUE  # If corresponding author is author 3, e.g., use CorrespAuthor_3: TRUE

# Comment out below to remove both. JEL Codes only given if keywords also given.
# keywords: "Multivariate GARCH \\sep Kalman Filter \\sep Copula" # Use \\sep to separate
# JELCodes: "L250 \\sep L100"

# ----- Manage headers and footers:
#BottomLFooter: $Title$
#BottomCFooter:
#TopLHeader: \leftmark # Adds section name at topleft. Remove comment to add it.
BottomRFooter: "\\footnotesize Page \\thepage" # Add a '#' before this line to remove footer.
addtoprule: TRUE
addfootrule: TRUE               # Use if footers added. Add '#' to remove line.

# --------- page margins:
margin: 2.3 # Sides
bottom: 2 # bottom
top: 2.5 # Top
HardSet_layout: TRUE # Hard-set the spacing of words in your document. This will stop LaTeX squashing text to fit on pages, e.g.
# This is done by hard-setting the spacing dimensions. Set to FALSE if you want LaTeX to optimize this for your paper.

# --------- Line numbers
linenumbers: FALSE # Used when submitting to journal

# ---------- References settings:
# You can download cls format here: https://www.zotero.org/ - simply search for your institution. You can also edit and save cls formats here: https://editor.citationstyles.org/about/
# Hit download, store it in Tex/ folder, and change reference below - easy.
bibliography: Tex/ref.bib       # Do not edit: Keep this naming convention and location.
csl: Tex/harvard-stellenbosch-university.csl # referencing format used.
# By default, the bibliography only displays the cited references. If you want to change this, you can comment out one of the following:
#nocite: '@*' # Add all items in bibliography, whether cited or not
# nocite: |  # add specific references that aren't cited
#  @grinold2000
#  @Someoneelse2010

# ---------- General:
RemovePreprintSubmittedTo: TRUE  # Removes the 'preprint submitted to...' at bottom of titlepage
Journal: "Journal of Finance"   # Journal that the paper will be submitting to, if RemovePreprintSubmittedTo is set to TRUE.
toc: FALSE                       # Add a table of contents
numbersections: TRUE             # Should sections (and thus figures and tables) be numbered?
fontsize: 11pt                  # Set fontsize
linestretch: 1.2                # Set distance between lines.
link-citations: FALSE            # This creates dynamic links to the papers in reference list.

### Adding additional latex packages:
header-includes:
#   - \usepackage{lscape}
#   - \usepackage{colortbl} # Add additional packages here.
#   - \setlength{\parskip}{0pt}      # Set the space between paragraphs to 0pt
#   - \setlength{\parindent}{1em}      # Set the indent size for the first line of paragraphs (adjust as needed)

output:
  pdf_document:
    keep_tex: TRUE
    template: Tex/TexDefault.txt
    fig_width: 3.5 # Adjust default figure sizes. This can also be done in the chunks of the text.
    fig_height: 3.5
abstract: 
---

<!-- First: Set your default preferences for chunk options: -->

<!-- If you want a chunk's code to be printed, set echo = TRUE. message = FALSE stops R printing ugly package loading details in your final paper too. I also suggest setting warning = FALSE and checking for warnings in R, else you might find ugly warnings in your paper. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
# Note: Include = FALSE implies the code is executed, but not printed in your pdf.
# warning and message = FALSE implies ugly messages and warnings are removed from your pdf.
# These should be picked up when you execute the command chunks (code sections below) in your rmd, not printed in your paper!
rm(list = ls())
# Lets load in example data, and see how this can be stored and later called from your 'data' folder.
source("code/install_and_load.R")
install_and_load(c("tidyverse","zoo","ggplot2","PerformanceAnalytics","lubridate","ggExtra","xts","tbl2xts","ggridges"))

list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))

ALSI <- read_rds("data/ALSI.rds")
ALSI$date <- as.Date(ALSI$date)
RebDays <- read_rds("data/Rebalance_days.rds")
Zar<-read_rds("data/Monthly_zar.rds")

```


<!-- ############################## -->
<!-- # Start Writing here: -->
<!-- ############################## -->

# Introduction \label{Introduction}

```{r distriution-by-vol-level}
# Example usage
plot_portfolio_returns_by_volatility(ALSI, Zar, compute_returns)
```


```{r plot-returns-by-cap-size}


# Example usage with a start date
ALSI %>% filter(!is.na(Index_Name)) %>% plot_mean_returns_by_cap_size(start_date = "2020-06-01")
```
Note that the returns sit on top of each other and therefore the plot is orange.  Only in the large caps do ALSI returns differ by generating a greater variance in mean returns.


```{r plot-returns-by-sector}


# Example usage with a start date
ALSI %>% filter(!is.na(Sector)) %>% plot_mean_returns_by_sector(start_date = "2020-06-01")
```


Again the returns sit on top of each other, and differ slightly when looking at Industrials and Resources sector. This indicates that holdings between of these sectors differ between the ALSI and SWIX.









```{r plot-sector-exposure-over-time}
ALSI %>%  plot_sector_exposure_over_time()

```
This graph shows that the SWIX tends to hold more Industrials.





```{r capping-impact-on-J403}
analyze_capped_indexes(ALSI, RebDays, column = "J403", caps = c(0.05, 0.10))

```


```{r capping-impact-on-J203}
analyze_capped_indexes(ALSI, RebDays, column = "J203", caps = c(0.05, 0.10))



```

Capping indexes reduce cumulative returns on both on the ALSI or SWIX.


