---
output:
  md_document:
    variant: markdown_github
---

# Question 5: Volatility and Garch estimates

I have been asked to evaluate the following two statements  

1. The South African rand (ZAR) has over the past few years been one of the most volatile currencies

2. The ZAR has generally performed well during periods where G10 currency carry trades have been favourable and currency valuations relatively cheap. Globally, it has been one of the currencies that most benefit during periods where the Dollar is comparatively strong, indicating a risk-on sentiment.


```{r,include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width = 6, fig.height = 5, fig.pos="H", fig.pos = 'H')
rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
source("code/install_and_load.R")
install_and_load(c("tidyverse","zoo","ggplot2","xts","treemap","PerformanceAnalytics","tbl2xts","rugarch","rmgarch","tidyr","scales"))
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))



cncy <- read_rds("data/currencies.rds")
cncy_Carry <- read_rds("data/cncy_Carry.rds")
cncy_value <- read_rds("data/cncy_value.rds")
cncyIV <- read_rds("data/cncyIV.rds")
bbdxy <- read_rds("data/bbdxy.rds")
IV <- read_rds("data/IV.rds")
```

```{r plot-sigma-comparisons,echo=FALSE}
# Prepare data for -sigZAR, including log returns in the result
zar_data <- cncy %>%
  filter(Name == "SouthAfrica_Cncy") %>%
  arrange(date) %>%
  mutate(log_return = log(Price / lag(Price))) %>%
  drop_na() %>%
  mutate(Squared_Returns = log_return^2)

# Convert to time-series format for GARCH model
zar_xts <- tbl_xts(zar_data, cols_to_xts = "log_return", spread_by = "Name")

# Fit GARCH model and extract conditional volatility
cond_vol_df <- fit_garch(zar_xts)


# Plot comparison
returns_df <- zar_data %>% select(date, Squared_Returns) %>% drop_na()
plot_volatility_comparison(returns_df, cond_vol_df, filter_date = "2010-01-01")

```






```{r plot-cond-var-for-multiple-countries,echo=FALSE}
# Define currencies
selected_currencies <- c(
  "SouthAfrica_Cncy", "Brazil_Cncy", "Australia_Cncy_Inv", "EU_Cncy_Inv", 
  "Japan_Cncy", "UK_Cncy_Inv", "Canada_Cncy", "Bostwana_Cncy_Inv", "China_Cncy"
)

# Fit GARCH for multiple currencies and pass to plotting funciton 
cncy %>% fit_garch_for_multiple_currencies( selected_currencies) %>% plot_volatility_across_currencies( highlight_currency = "SouthAfrica_Cncy", filter_date = "2015-01-01")

```
The graph confirms that the South African Rand is highly volatile compared to other major currencies.



```{r plot-dynamic-correlations,echo=FALSE}

# Calculate log returns for ZAR and Dollar
zar_data <- calculate_log_returns(cncy %>% filter(Name == "SouthAfrica_Cncy"))
dollar_data <- calculate_log_returns(bbdxy %>% filter(Name == "BBDXY"))

# Prepare PCA data
pc_data <- prepare_pca_data(IV %>% filter(Name %in% c("V2X", "VIX", "VXEEM")))

# Merge ZAR, Dollar, and PC1 data
merged_data <- zar_data %>%
  select(date, ZAR_log_return = log_return) %>%
  left_join(dollar_data %>% select(date, Dollar_log_return = log_return), by = "date") %>%
  left_join(pc_data %>% select(date, PC1), by = "date") %>%
  drop_na()

# Fit DCC-GARCH model
dcc_fit <- fit_dcc_garch(merged_data)

# Prepare data frames for plotting
volatility_df <- prepare_volatility_df(dcc_fit, merged_data$date)
correlation_df <- prepare_correlation_df(dcc_fit, merged_data$date)

plot_dynamic_correlations(correlation_df)


```

The graph partially supports the second statment. It shows that ZAR correlates positively with the Dollar (which suggests ZAR benefits in certain global conditions) and responds inversely to volatility. This indicates that ZAR may perform well during periods when global markets favour risk-taking, but it may not be straightforwardly linked to the Dollar's strength in all cases.



