---
output:
  md_document:
    variant: markdown_github
---

# Question 5: Volatility and Garch estimates

I have been asked to evaluate the following two statements  

1. The South African rand (ZAR) has over the past few years been one of the most volatile currencies

2. The ZAR has generally performed well during periods where G10 currency carry trades have been favourable and currency valuations relatively cheap. Globally, it has been one of the currencies that most benefit during periods where the Dollar is comparatively strong, indicating a risk-on sentiment.


```{r,include=FALSE}

rm(list = ls()) # Clean your environment:
gc() # garbage collection - It can be useful to call gc after a large object has been removed, as this may prompt R to return memory to the operating system.
source("code/install_and_load.R")
install_and_load(c("tidyverse","zoo","ggplot2","xts","treemap","PerformanceAnalytics","tbl2xts","rugarch","rmgarch","tidyr","scales"))
list.files('code/', full.names = T, recursive = T) %>% .[grepl('.R', .)] %>% as.list() %>% walk(~source(.))



cncy <- read_rds("data/currencies.rds")
cncy_Carry <- read_rds("data/cncy_Carry.rds")
cncy_value <- read_rds("data/cncy_value.rds")
cncyIV <- read_rds("data/cncyIV.rds")
bbdxy <- read_rds("data/bbdxy.rds")
IV <- read_rds("data/IV.rds")
```

```{r plot-sigma-comparisons,echo=FALSE}
# Prepare data for -sigZAR, including log returns in the result
zar_data <- cncy %>%
  filter(Name == "SouthAfrica_Cncy") %>%
  arrange(date) %>%
  mutate(log_return = log(Price / lag(Price))) %>%
  drop_na() %>%
  mutate(Squared_Returns = log_return^2)

# Convert to time-series format for GARCH model
zar_xts <- tbl_xts(zar_data, cols_to_xts = "log_return", spread_by = "Name")

# Fit GARCH model and extract conditional volatility
cond_vol_df <- fit_garch(zar_xts)


# Plot comparison
returns_df <- zar_data %>% select(date, Squared_Returns) %>% drop_na()
plot_volatility_comparison(returns_df, cond_vol_df, filter_date = "2010-01-01")

```

 This figure directly illustrates the core concept of GARCH models: capturing time-varying volatility. By comparing the historical volatility (Returns Sigma) with the GARCH-predicted volatility (Sigma from GARCH), you can visually demonstrate:

Volatility Clustering: GARCH models are particularly adept at capturing periods of high and low volatility, which are often clustered together. Your figure can highlight these clusters.
Model Fit: The closer the two lines are, the better the GARCH model fits the data. Deviations between the lines can indicate periods where the model under- or overestimates volatility.
Forecasting Power: If you've used the GARCH model to generate forecasts, you can include future predictions on the plot. This will show how well the model anticipates future volatility.




```{r plot-cond-var-for-multiple-countries,echo=FALSE}
# Define currencies
selected_currencies <- c(
  "SouthAfrica_Cncy", "Brazil_Cncy", "Australia_Cncy_Inv", "EU_Cncy_Inv", 
  "Japan_Cncy", "UK_Cncy_Inv", "Canada_Cncy", "Bostwana_Cncy_Inv", "China_Cncy"
)

# Fit GARCH for multiple currencies and pass to plotting funciton 
cncy %>% fit_garch_for_multiple_currencies( selected_currencies) %>% plot_volatility_across_currencies( highlight_currency = "SouthAfrica_Cncy", filter_date = "2015-01-01")

```




```{r plot-dynamic-correlations,echo=FALSE}

# Calculate log returns for ZAR and Dollar
zar_data <- calculate_log_returns(cncy %>% filter(Name == "SouthAfrica_Cncy"))
dollar_data <- calculate_log_returns(bbdxy %>% filter(Name == "BBDXY"))

# Prepare PCA data
pc_data <- prepare_pca_data(IV %>% filter(Name %in% c("V2X", "VIX", "VXEEM")))

# Merge ZAR, Dollar, and PC1 data
merged_data <- zar_data %>%
  select(date, ZAR_log_return = log_return) %>%
  left_join(dollar_data %>% select(date, Dollar_log_return = log_return), by = "date") %>%
  left_join(pc_data %>% select(date, PC1), by = "date") %>%
  drop_na()

# Fit DCC-GARCH model
dcc_fit <- fit_dcc_garch(merged_data)

# Prepare data frames for plotting
volatility_df <- prepare_volatility_df(dcc_fit, merged_data$date)
correlation_df <- prepare_correlation_df(dcc_fit, merged_data$date)

plot_dynamic_correlations(correlation_df)


```

The graph partially supports the second statment. It shows that ZAR correlates positively with the Dollar (which suggests ZAR benefits in certain global conditions) and responds inversely to volatility. This indicates that ZAR may perform well during periods when global markets favour risk-taking, but it may not be straightforwardly linked to the Dollar's strength in all cases.





























works










































<!-- ```{r} -->
<!-- library(dplyr) -->
<!-- library(tidyr) -->
<!-- library(stats) -->

<!-- zar_data <- cncy %>% -->
<!--   filter(Name == "SouthAfrica_Cncy") %>% -->
<!--   arrange(date) %>% -->
<!--   mutate(log_return = log(Price / lag(Price))) %>% -->
<!--   drop_na() -->




<!-- # Step 3: Extract Dollar index data (BBDXY) -->
<!-- dollar_data <- bbdxy %>% -->
<!--   filter(Name == "BBDXY") %>% -->
<!--   arrange(date) %>% -->
<!--   mutate(log_return = log(Price / lag(Price))) %>% -->
<!--   drop_na() -->


<!-- # Compute log returns for volatility indices -->
<!-- pc_data <- IV %>% -->
<!--   filter(Name %in% c("V2X", "VIX", "VXEEM")) %>% -->
<!--   pivot_wider(names_from = Name, values_from = Price) %>% -->
<!--   mutate(across(starts_with("V"), ~ log(. / lag(.)))) %>% -->
<!--   drop_na() -->

<!-- # Perform PCA -->
<!-- pc_result <- prcomp(pc_data %>% select(starts_with("V")), scale. = TRUE) -->

<!-- # Extract PC1 -->
<!-- pc_data <- pc_data %>% -->
<!--   mutate(PC1 = predict(pc_result)[, 1]) -->


<!-- # Merge ZAR and Dollar log returns with PC1 -->
<!-- merged_data <- zar_data %>% -->
<!--   select(date, ZAR_log_return = log_return) %>% -->
<!--   left_join(dollar_data %>% select(date, Dollar_log_return = log_return), by = "date") %>% -->
<!--   left_join(pc_data %>% select(date, PC1), by = "date") %>% -->
<!--   drop_na() -->


<!-- ``` -->




<!-- ```{r} -->
<!-- library(rugarch) -->
<!-- library(rmgarch) -->

<!-- # Univariate GARCH specification -->
<!-- univariate_spec <- ugarchspec( -->
<!--   variance.model = list(model = "gjrGARCH", garchOrder = c(1, 1)), -->
<!--   mean.model = list(armaOrder = c(0, 0), include.mean = TRUE), -->
<!--   distribution.model = "std" -->
<!-- ) -->

<!-- # Multivariate GARCH specification -->
<!-- multi_univ_garch_spec <- multispec(replicate(ncol(merged_data) - 1, univariate_spec)) # Exclude date -->
<!-- dcc_spec <- dccspec( -->
<!--   uspec = multi_univ_garch_spec, -->
<!--   dccOrder = c(1, 1), -->
<!--   distribution = "mvnorm" -->
<!-- ) -->

<!-- # Fit the model -->
<!-- dcc_fit <- dccfit(dcc_spec, data = as.matrix(merged_data[, -1]))  # Exclude date -->

<!-- ``` -->


<!-- ```{r} -->
<!-- volatilities <- sigma(dcc_fit) -->
<!-- head(volatilities) -->

<!-- ``` -->

<!-- ```{r} -->
<!-- correlations <- rcor(dcc_fit) -->
<!-- head(correlations[1, 2, ])  # Correlation between ZAR and Dollar -->
<!-- head(correlations[1, 3, ])  # Correlation between ZAR and PC1 -->

<!-- ``` -->


<!-- ```{r} -->
<!-- print(dcc_fit) -->

<!-- ``` -->



<!-- ```{r} -->
<!-- # Ensure volatilities have the correct date attached -->
<!-- volatility_df <- data.frame( -->
<!--   date = merged_data$date, -->
<!--   ZAR_volatility = volatilities[, 1], -->
<!--   Dollar_volatility = volatilities[, 2], -->
<!--   PC1_volatility = volatilities[, 3] -->
<!-- ) %>% -->
<!--   pivot_longer(-date, names_to = "Series", values_to = "Volatility") -->

<!-- ``` -->





<!-- ```{r} -->
<!-- # Ensure correlations have the correct date attached -->
<!-- correlation_df <- data.frame( -->
<!--   date = merged_data$date, -->
<!--   ZAR_Dollar = correlations[1, 2, ], -->
<!--   ZAR_PC1 = correlations[1, 3, ] -->
<!-- ) %>% -->
<!--   pivot_longer(-date, names_to = "Correlation_Pair", values_to = "Correlation") -->

<!-- ``` -->



<!-- ```{r} -->


<!-- # Plot -->
<!-- ggplot(correlation_df, aes(x = date, y = Correlation, color = Correlation_Pair)) + -->
<!--   geom_line() + -->
<!--   labs( -->
<!--     title = "Dynamic Correlations over Time", -->
<!--     x = "Date", -->
<!--     y = "Correlation", -->
<!--     color = "Pair" -->
<!--   ) + -->
<!--   theme_minimal() -->

<!-- ``` -->


<!-- The graph partially supports the statement. It shows that ZAR correlates positively with the Dollar (which suggests ZAR benefits in certain global conditions) and responds inversely to volatility. This indicates that ZAR may perform well during periods when global markets favor risk-taking, but it may not be straightforwardly linked to the Dollar's strength in all cases. -->





<!-- ```{r} -->
<!-- plot(dcc_fit, which = 2)  # Residuals plot -->
<!-- plot(dcc_fit, which = 3)  # ACF plot of residuals -->
<!-- plot(dcc_fit, which = 4)  # PACF plot of residuals -->

<!-- ``` -->





























<!-- ```{r} -->
<!-- library(ggplot2) -->
<!-- library(tidyr) -->

<!-- # Convert volatilities to a tidy format -->
<!-- vol_data <- data.frame(date = merged_data$date, -->
<!--                        ZAR_volatility = volatilities[, 1], -->
<!--                        Dollar_volatility = volatilities[, 2], -->
<!--                        PC1_volatility = volatilities[, 3]) -->

<!-- # Tidy data for ggplot -->
<!-- vol_data_tidy <- vol_data %>% -->
<!--   gather(key = "Variable", value = "Volatility", -date) -->

<!-- # Plot volatilities -->
<!-- ggplot(vol_data_tidy, aes(x = date, y = Volatility, color = Variable)) + -->
<!--   geom_line() + -->
<!--   theme_minimal() + -->
<!--   labs(title = "Conditional Volatilities",  -->
<!--        x = "Date",  -->
<!--        y = "Volatility") + -->
<!--   theme(legend.position = "bottom") + -->
<!--   scale_color_manual(values = c("blue", "red", "green")) -->

<!-- ``` -->




















<!-- ```{r} -->
<!-- rtn <- merged_data %>% -->
<!--   mutate(across(starts_with("log_return"), ~ . - mean(., na.rm = TRUE))) %>% -->
<!--   drop_na() -->

<!-- library(rugarch) -->
<!-- univariate_spec <- ugarchspec( -->
<!--   variance.model = list(model = "gjrGARCH", garchOrder = c(1, 1)), -->
<!--   mean.model = list(armaOrder = c(0, 0), include.mean = TRUE), -->
<!--   distribution.model = "std"  # Student-t distribution -->
<!-- ) -->

<!-- fits <- lapply(rtn %>% select(starts_with("log_return")), function(series) { -->
<!--   ugarchfit(spec = univariate_spec, data = series) -->
<!-- }) -->

<!-- lapply(fits, function(fit) fit@fit$convergence) -->

<!-- library(rmgarch) -->
<!-- multi_univ_garch_spec <- multispec(replicate(ncol(rtn), univariate_spec)) -->
<!-- dcc_spec <- dccspec( -->
<!--   uspec = multi_univ_garch_spec, -->
<!--   dccOrder = c(1, 1), -->
<!--   distribution = "mvnorm" -->
<!-- ) -->


<!-- dcc_fit <- dccfit(dcc_spec, data = as.matrix(rtn), fit.control = list(eval.se = TRUE)) -->

<!-- volatilities <- sigma(dcc_fit) -->
<!-- correlations <- rcor(dcc_fit) -->


<!-- plot(volatilities) -->
<!-- plot(correlations[1, 2, ])  # Correlation between ZAR and another series -->


<!-- diag_results <- MCHdiag(rtn, rcov(dcc_fit)) -->

<!-- ``` -->





<!-- ```{r} -->
<!-- library(dplyr) -->
<!-- library(tidyr) -->

<!-- # Step 1: Extract ZAR data from cncy -->
<!-- zar_data <- cncy %>% -->
<!--   filter(Name == "SouthAfrica_Cncy") %>% -->
<!--   arrange(date) %>% -->
<!--   mutate(log_return = log(Price / lag(Price))) %>% -->
<!--   drop_na() -->

<!-- # Step 2: Extract G10 carry trade data (DBHVG10U) -->
<!-- carry_data <- cncy_Carry %>% -->
<!--   filter(Name == "DBHVG10U") %>% -->
<!--   arrange(date) %>% -->
<!--   mutate(log_return = log(Price / lag(Price))) %>% -->
<!--   drop_na() -->

<!-- # Step 3: Extract Dollar index data (BBDXY) -->
<!-- dollar_data <- bbdxy %>% -->
<!--   filter(Name == "BBDXY") %>% -->
<!--   arrange(date) %>% -->
<!--   mutate(log_return = log(Price / lag(Price))) %>% -->
<!--   drop_na() -->

<!-- # Step 4: Merge all data on the common 'date' column -->
<!-- merged_data <- zar_data %>% -->
<!--   select(date, ZAR_log_return = log_return) %>% -->
<!--   left_join(carry_data %>% -->
<!--               select(date, Carry_log_return = log_return), by = "date") %>% -->
<!--   left_join(dollar_data %>% -->
<!--               select(date, Dollar_log_return = log_return), by = "date") %>% -->
<!--   drop_na()  # Remove any rows with missing values after the join -->

<!-- # View the first few rows of the merged data -->
<!-- # univariate_spec <- ugarchspec( -->
<!-- #   variance.model = list(model = "sGARCH", garchOrder = c(1, 1)), -->
<!-- #   mean.model = list(armaOrder = c(0, 0), include.mean = TRUE), -->
<!-- #   distribution.model = "norm" -->
<!-- # ) -->



<!-- # Step 3: Create the multivariate GARCH specification (DCC-GARCH) -->
<!-- dcc_spec <- dccspec( -->
<!--   uspec = multispec(replicate(ncol(merged_data) - 1, univariate_spec)),  # Remove the 'date' column -->
<!--   dccOrder = c(1, 1), -->
<!--   distribution = "mvnorm" -->
<!-- ) -->

<!-- # Step 4: Fit the DCC-GARCH model -->
<!-- dcc_fit <- dccfit(dcc_spec, data = as.matrix(merged_data[, -1]))  # Remove 'date' column for fitting -->

<!-- # Step 5: Extract results (conditional volatilities and correlations) -->
<!-- dcc_volatilities <- sigma(dcc_fit)  # Extract conditional volatilities -->
<!-- dcc_correlations <- rcor(dcc_fit)   # Extract conditional correlations -->

<!-- # Step 6: Display the results (first few rows) -->
<!-- head(dcc_volatilities) -->
<!-- head(dcc_correlations) -->

<!-- ``` -->





<!-- ```{r} -->
<!-- library(rugarch) -->
<!-- library(rmgarch) -->

<!-- # Prepare data for multivariate GARCH -->
<!-- merged_data <- cncy %>% -->
<!--   filter(Name %in% c("SouthAfrica_Cncy", "DBHVG10U", "BBDXY")) %>% -->
<!--   arrange(date) %>% -->
<!--   group_by(Name) %>% -->
<!--   mutate(log_return = log(Price / lag(Price))) %>% -->
<!--   ungroup() %>% -->
<!--   drop_na() %>% -->
<!--   pivot_wider(names_from = Name, values_from = log_return) -->

<!-- # Convert to time-series matrix -->
<!-- returns_matrix <- as.matrix(merged_data[, -1]) -->

<!-- # Specify univariate GARCH for each series -->
<!-- univariate_spec <- ugarchspec( -->
<!--   variance.model = list(model = "sGARCH", garchOrder = c(1, 1)), -->
<!--   mean.model = list(armaOrder = c(0, 0), include.mean = TRUE), -->
<!--   distribution.model = "norm" -->
<!-- ) -->

<!-- # Specify DCC-GARCH model -->
<!-- dcc_spec <- dccspec( -->
<!--   uspec = multispec(replicate(ncol(returns_matrix), univariate_spec)), -->
<!--   dccOrder = c(1, 1), -->
<!--   distribution = "mvnorm" -->
<!-- ) -->

<!-- # Fit the model -->
<!-- dcc_fit <- dccfit(dcc_spec, data = returns_matrix) -->

<!-- # Extract conditional correlations and volatilities -->
<!-- dcc_correlations <- rcor(dcc_fit)  # Conditional correlations -->
<!-- dcc_volatilities <- sigma(dcc_fit)  # Conditional volatilities -->

<!-- ``` -->




























